---
title: "R Bindings for Netica :: Cheat Sheet"
description: " "
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin

source("common.R")
use_cheatsheet_logo(
  "Peanut", 
  alt = "Hex logo for Peanut - A peanut with a small three node network."
)

sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

<!-- Page 1 -->

## Netica Licence Keys

The Netica shared library which is the heart of RNetica is availble
under license from Norsys, LLC (<https://www.norsys.com>).  Unlocking
its full potential requires purchasing an API (not GUI) licences key
from Norsys.  This will be a string.  There are several different
places you can set this.

- Create an environment variable, `NETICA_LICENSE_KEY` in `.Renviorn`
  or other place.
- Set `options("NeticaLicenseKey")` to the string.
- Pass it in explicitly with a call to `NeticaSession`.

Sessions started without a licence key will still operate, but the
size the networks will be limited.

## Netica Naming conventions

Netica networks, nodes, and states must have a name that (1) start
with a letter, (2) have 30 characters or fewer, and (3) all characters
are alphanumeric or `_`.  While names must conform to the Netica
conventions, titles do not.

	- `r args(is.IDname)`, `r args(as.IDname)`

## Netica Sessions

To use RNetica, a `NeticaSession` object must be created and started.
Networks and other objects are then created within a session.  A
session, network or node object is active if it is connected to an
object inside of the Netica session.  When
the session is stoped (including when R exits) the connection to all
networks and nodes is lost and they become inactive.

	- `r args(NeticaSession)` -- Generateds a new session
    - `r args(startSession)`, `r args(stopSession)`
	- `r args(is.active)` -- Tests connection of object to Netica session.
	- `r args(getDefaultSession)` -- Returns the default value used in
	function which expect a session.
	- `r args(GetNthNetwork)` -- Used for looping through open
	networks, returns `NULL` when all networks processed.
	- `r args(GetNamedNeworks)` -- Returns networks by name.
	- `session$nets$<netname>` -- The `$nets` field of the session
	argument stores the network objects.  `ls(session$nets)` lists them.
	- `r args(CheckNamedNetworks)` -- Searches Netica memory for
	matching network.
	
## Error Handling	



- `NeticaCondition`
- `r args(ReportErrors)`
- `r args(ClearAllErrors)`
- `r args(logErrors)`, `r args(flogErrors)`

## Networks

- `NeticaBN`
-  is.NeticaBN,
- `r args(NetworkName)`, `r args(NetworkTitle)`, `r args(NetworkComment)`
-  CreateNetwork, DeleteNetwork, CopyNetworks
- `r args(ReadNetworks)`, `r args(WriteNetworks)`
- `r args(GetNetworkFileName)`
-  NetworkUserField, 
-  NetworkUserObj, 
-  NetworkAllUserFields
-  NetworkSession
-  NetworkAllNodes,
-  `net$nodes$<nodename>`
-  NetworkFindNode,

## Temporary network creation

	- `r args(SetupDefaultSession)`
   local_create_nets,local_load_nets,
       local_copy_nets,local_RNetica_net


## Nodes

  NewDiscreteNode, NewContinuousNode, DeleteNodes,
  CopyNodes,
  NodeNet, NodeName, "NodeName<-",
  is.NeticaNode,
  NodeTitle, "NodeTitle<-",
  NodeDescription, "NodeDescription<-",
  NodeUserField, "NodeUserField<-", NodeAllUserFields,
  NodeUserObj, "NodeUserObj<-",
  NodeKind, "NodeKind<-",
  NodeVisStyle, "NodeVisStyle<-",
  NodeVisPos, "NodeVisPos<-"

## Node States

  is.discrete, is.continuous,
  NodeStates, NodeNumStates, "NodeStates<-",
  NodeStateTitles, "NodeStateTitles<-",
  NodeStateComments, "NodeStateComments<-",
  NodeLevels, "NodeLevels<-", ParentStates, ParentNames


## Edges

- `CliqueNode`
   AddLink, DeleteLink, ReverseLink,
   NodeParents, "NodeParents<-", NodeChildren,
   AbsorbNodes, MakeCliqueNode, is.CliqueNode,
   GetClique,
   is.NodeRelated, GetRelatedNodes,
   NodeInputNames, "NodeInputNames<-",
   AdjoinNetwork, NetworkFootprint

## CPTs

- `[`, `[[`
-   NodeProbs, "NodeProbs<-", EVERY_STATE,
  IsNodeDeterministic, HasNodeTable,
  DeleteNodeTable
    NodeEquation, "NodeEquation<-",EquationToTable,


## Inference

   CompileNetwork, UncompileNetwork, is.NetworkCompiled,
   EliminationOrder, "EliminationOrder<-",
   NetworkCompiledSize, JunctionTreeReport,
   NodeFinding, "NodeFinding<-", EnterNegativeFinding,
   RetractNetFindings, RetractNodeFinding, EnterFindings,
   NodeBeliefs, IsBeliefUpdated, NodeMode,
   NodeLikelihood, "NodeLikelihood<-",
   JointProbability, FindingsProbability, MostProbableConfig
   GetNetworkAutoUpdate, SetNetworkAutoUpdate, WithoutAutoUpdate,
   NodeValue, "NodeValue<-",
   EnterGaussianFinding, EnterIntervalFinding,
   NodeExpectedValue, NodeExpectedUtils,
   CalcNodeState, CalcNodeValue,
   MutualInfo, VarianceOfReal,
   woe, ewoe


## Node Sets

    NodeSets, "NodeSets<-",
    AddNodeToSets, RemoveNodeFromSets,
    NetworkNodeSets, NetworkSetPriority,
    NetworkNodesInSet, "NetworkNodesInSet<-",
    NetworkNodeSetColor

## Cases
	- `CaseStream`, `FileCaseStream`, `MemoryCaseStream`
	   CaseFileDelimiter,CaseFileMissingCode,
       OpenCaseStream, CloseCaseStream,
       WithOpenCaseStream,
       is.NeticaCaseStream, isCaseStreamOpen,
       CaseMemoryStream, CaseFileStream,
       is.MemoryCaseStream, is.CaseFileStream,
       getCaseStreamPath,getCaseStreamLastId,
       getCaseStreamPos,getCaseStreamLastFreq,
       getCaseStreamDataFrameName,
       MemoryStreamContents,"MemoryStreamContents<-",
       read.CaseFile,write.CaseFile,orderVars,
       WriteFindings,ReadFindings,
       NodeExperience, "NodeExperience<-",
       FadeCPT, LearnFindings,
       LearnCases, LearnCPTs


## Random Numbers

GenerateRandomCase, NetworkSetRNG, WithRNG, isNeticaRNGActive,
       is.NeticaRNG, FreeNeticaRNG, NewNeticaRNG
	   
## Tester

testerNet, testerTarget, testerIgnore,
       testerErrorRate, testerLogLoss, testerQuadraticLoss,
       testerConfusion, testerKappa, testerLambda,
       testNetwork


## Utilities

       dputToString, dgetFromString





------------------------------------------------------------------------

CC BY SA Russell G. Almond • [ralmond\@fsu.edu](mailto:ralmond@fsu.edu) • [ralmond.net](https://ralmond.net)

Learn more at [https://ralmond.net/Peanut](https://ralmond.net/Peanut).

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true

packageVersion("RNetica")
```

------------------------------------------------------------------------
