---
title: "Simulation Studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SimulationStudies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RNetica)
library(CPTtools)
```

## Netica Setup

If you have a license key from Norsys, set the value of `"NETICA_LICENSE_KEY"` in `~/.Renviron`.

```{r startNetica}
sess <- NeticaSession(LicenseKey = Sys.getenv("NETICA_LICENSE_KEY"))
startSession(sess)
```

## Build the network for the test form.

This exercise uses the language assessment proposed in @Mislevy1995.

Based on that paper, the following Netica files were built (and are distributed in the "sampleNets" directory of the `RNetica` package.)

| Model       | Filename          | Variables                             | Replications |
|----------------|--------------------------|----------------|----------------|
| Proficiency | LanguagePM.dne    | Reading, Writing, Speaking, Listening | `NA`         |
| Reading     | ReadingTask.dne   | TaskA                                 | 5            |
| Writing     | WritingTask.dne   | TaskB                                 | 3            |
| Speaking    | SpeakingTask.dne  | TaskC                                 | 3            |
| Listening   | ListeningTask.dne | TaskD                                 | 5            |

: Networks in Language Model

```{r form design}

pmname <- "LanguagePM"
tasks <- c("Reading","Writing","Speaking","Listening")
reps.form16<-c(5,3,3,5)
```

Load the proficiency and evidence models.

```{r LoadNets}
pmfile <- system.file(paste0("sampleNets/",
                             pmname,".dne"),
                      package="RNetica")
pm <- ReadNetworks(pmfile,session=sess)

emfiles <- system.file(paste0("sampleNets/",
                              tasks,
                              "Task.dne"),
                       package="RNetica")
ems <- ReadNetworks(emfiles,session=sess)
```

Putting the proficiency variables and observable variables into node sets will make it easier to find them later.

```{r nodeSets}
proficiencies <- NetworkAllNodes(pm)
NetworkNodesInSet(pm,"Proficiencies") <- proficiencies
for (em in ems) {
  NetworkNodesInSet(em,"Observables") <- NetworkAllNodes(em)
}
```

## Build the motif

While the proficiency model is a complete Bayesian network, the evidence models (with "Task" in their name) are fragments. In particular, they have unconnected edges corresponding to the node of the proficiency variables. The `AdjoinNetwork()` function copies the evidence model into the active network and reconnects the edges. A network containing the proficiency model plus one or more evidence models is a *motif*.

For testing purposes, motifs with just the proficiency variable and the evidence model for a single task are useful.

For evaluating an assessment form, a motif consisting of the proficiency model plus evidence models for all of the tasks is what is needed. To construct this, adjoin copies of the evidence models according the the form design given in the `reps` variable.

```{r buildMotif}
## If reruning this code, delete the motif.
if (exists("motif16") && is.active(motif16))
  DeleteNetwork(motif16)
motif16 <- CopyNetworks(pm,"motif16")
for (whichTask in 1:length(reps.form16)) {
  for (iTasks in 1:reps.form16[whichTask]) {
    newobs <- AdjoinNetwork(motif16,ems[[whichTask]])
  }
}
observables <- NetworkNodesInSet(motif16,"Observables")
names(observables)
```

```{r profs}
proficiencies <- NetworkNodesInSet(motif16,"Proficiencies")
names(proficiencies)
```

Finally, compile the network.

```{r compile}
CompileNetwork(motif16)
```

# The Simulation

First set some constants. `B` is the number of cases to simulate.

```{r simData}
B <- 1000L
casefile <- tempfile("motif16test",fileext=".cas")
filestream <- CaseFileStream(casefile, session=sess)
rng <- NewNeticaRNG(123456779, session=sess)
allvars <- c(proficiencies,observables)
```

## Setup space for the simulation.

The sample data and true values for the proficiency variables. However, need a space for the margin scores. Pre-allocating the space makes the simulation run faster.

```{r preallocation}
varstatenames <- 
  names(unlist(lapply(proficiencies,NodeBeliefs)))
motif16.margins <- matrix(NA_real_,B,length(varstatenames))
colnames(motif16.margins) <- varstatenames
motif16.modes <- matrix(NA_character_,B,length(proficiencies))
colnames(motif16.modes) <- names(proficiencies)
```

## Simulate Proficiency Variables and observables

Loop through the simulees.

1.  Generate the data using the motif.

2.  Write out the cases.

3.  Retract the generated proficiency variables.

4.  Collect the marginal distributions for the proficiency variables.

5.  Unset all variables.

```{r}
WithOpenCaseStream(filestream,
  WithRNG(rng,
    for (b in 1L:B) {
      ## 1.  Generate the data
      GenerateRandomCase(allvars,rng=rng)
      ## 2. Write the findings both proficiency and observable variables.
      WriteFindings(allvars,filestream,b)
      ## 3. Retract the proficiency "findings", this leaves the marginal "scores" of the remaining variables.
      lapply(proficiencies,RetractNodeFinding) 
      ## 4. Save the marginal distributions
      motif16.margins[b,] <- unlist(lapply(proficiencies,NodeBeliefs))
      motif16.modes[b,] <- unlist(lapply(proficiencies,NodeMode))
      ## 5. Now clear out the rest of the nodes.
      RetractNetFindings(motif16)
    }))
```

Read the case data back in, and bind the raw data to the scores. Use the `orderVars` function to convert to factors.

```{r readCaseData}
motif16.data <- orderVars(read.CaseFile(casefile,session=sess),
                          allvars)
motif16.modes <- orderVars(as.data.frame(motif16.modes),proficiencies)
motif16.data <- data.frame(motif16.data,motif16.margins,
                           mode=motif16.modes)
head(motif16.data)
```

# Evaluate the Form

## Build Confusion Matrixes

```{r}
motif16.cm <- 
  lapply(names(proficiencies), function (p)
         table(motif16.data[,paste0(c("","mode."),p)]))
names(motif16.cm) <- names(proficiencies)
motif16.cm$Reading
```

```{r}
lapply(motif16.cm,gkLambda)
```

```{r}
lapply(motif16.cm,fcKappa)
```



## Build Expected Confusion Matrixes

```{r}
motif16.em <- 
  lapply(names(proficiencies), function (p)
         expTable(motif16.data,p,p,"<var>\\.<state>"))
names(motif16.em) <- names(proficiencies)
motif16.em$Reading
```

## Goodman and Kruskal Lambda

## Cohen's Kappa

# Try again with a longer test:

Now use six tasks of each type.

```{r buildMotif24}
motif24.addreps <- c(1,3,3,1)
reps.form24 <- reps.form16+motif24.addreps
## If reruning this code, delete the motif.
if (exists("motif24") && is.active(motif24))
  DeleteNetwork(motif24)
motif24 <- CopyNetworks(motif16,"motif24")
for (whichTask in 1:length(motif24.addreps)) {
  for (iTasks in 1:motif24.addreps[whichTask]) {
    newobs <- AdjoinNetwork(motif24,ems[[whichTask]])
  }
}
observables <- NetworkNodesInSet(motif24,"Observables")
## Same names, but different nodes in different networks.
proficiencies <- NetworkNodesInSet(motif24,"Proficiencies")
CompileNetwork(motif24)
```

## Redo the simulation

```{r simData24}
B <- 1000L
casefile <- tempfile("motif24test",fileext=".cas")
filestream <- CaseFileStream(casefile, session=sess)
allvars <- c(proficiencies,observables)
rng <- NewNeticaRNG(677912356, session=sess)
```

```{r preallocation24}
varstatenames <- 
  names(unlist(lapply(proficiencies,NodeBeliefs)))
motif24.margins <- matrix(NA_real_,B,length(varstatenames))
colnames(motif24.margins) <- varstatenames
motif24.modes <- matrix(NA_character_,B,length(proficiencies))
colnames(motif24.modes) <- names(proficiencies)
```

```{r}
WithOpenCaseStream(filestream,
  WithRNG(rng,
    for (b in 1L:B) {
      ## 1.  Generate the data
      GenerateRandomCase(allvars,rng=rng)
      ## 2. Write the findings both proficiency and observable variables.
      WriteFindings(allvars,filestream,b)
      ## 3. Retract the proficiency "findings", this leaves the marginal "scores" of the remaining variables.
      lapply(proficiencies,RetractNodeFinding) 
      ## 4. Save the marginal distributions
      motif24.margins[b,] <- unlist(lapply(proficiencies,NodeBeliefs))
      motif24.modes[b,] <- unlist(lapply(proficiencies,NodeMode))
      ## 5. Now clear out the rest of the nodes.
      RetractNetFindings(motif24)
    }))
```

```{r readCaseData24}
motif24.data <- orderVars(read.CaseFile(casefile,session=sess),
                          allvars)
motif24.modes <- orderVars(as.data.frame(motif16.modes),proficiencies)
motif24.data <- data.frame(motif24.data,motif16.margins,
                           mode=motif24.modes)
head(motif24.data)
```
